---
title: "Turkish Energy Policy Analysis"
author: "Ahmet Şamil Tür, Didem Ateş, Güneş Uğuz, Irmak Bolat, İrem Kaya"
subtitle: "Econ 413 Term Project"
output: html_document
---

<style type="text/css">
body {
  font-family: Calibri ;
  font-size: 14pt;
  color:black;
}
h1,h2,h3 {
  font-family: Calibri ;
  font-size: 18pt;
  color: darkblue;
}
h4,h5,h6 {
  font-family: Calibri ;
  font-size: 16pt;
  color: darkblue;
}
.tabset {
font-family: Calibri ;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, results = F)
```

```{r run_scripts, results=T}
source("R/01_download_pdf.R", encoding = "UTF-8")
source("R/02_read_pdf.R")
```

```{r search_keywords}

search_terms <- c("yenilenebilir", "rüzgar", "rüzgâr", "güneş", "nükleer",
                  "hidroelektrik", "jeotermal", "ngs", "hes", "petrol", "gaz", 
                  "kömür", "doğalgaz", "linyit")

freqs <- findFreq(search_terms)

freqsDT <- as.data.table(freqs, keep.rownames = "Years")
freqsDT$gaz <- freqsDT$gaz + freqsDT$doğalgaz
freqsDT$hidroelektrik <- freqsDT$hidroelektrik + freqsDT$hes
freqsDT$nükleer <- freqsDT$nükleer + freqsDT$ngs
freqsDT$kömür <- freqsDT$kömür + freqsDT$linyit
# sums[search_terms]

# findFreq("enerji")
# sums["enerji"]

# # all words starting with verim Adjust regex to find what you need.
# freq[,grep("^verim", colnames(freq))]
# sums[grep("^verim", names(sums))]
```

```{r libraries}
my_packages <- c("WDI", "data.table", "ggplot2", "plotly", "ggthemes", "scales")
not_installed <- my_packages[!(my_packages %in% installed.packages()[ , "Package"])]
if(length(not_installed)) install.packages(not_installed, repos = "http://cran.us.r-project.org")

library(WDI)
library(data.table)
library(ggplot2)
library(plotly)
library(ggthemes)
library(scales)
```

```{r theme}
my_theme <- function()  { theme_bw()   +
    theme(panel.border = element_rect(color = "gray70", size = 1.5)) +
  theme(legend.position = "none") +
  theme(plot.caption = element_text(color = "blue4", size = rel(0.9))) +
  theme(panel.background = element_rect(fill = "mintcream", colour = "gray70"), panel.grid = element_line(colour ="gray70"),
      panel.grid.minor = element_line(size = rel(0.5)),
      strip.background = element_rect(fill = "mintcream",
                              colour = "gray70"), complete = TRUE) +theme(axis.title.x= element_text(color = "blue4", size = rel(0.9)),
                            axis.title.y = element_text(color = "blue4", size = rel(0.9))) +
    theme(plot.title = element_text(color = "blue4", size = rel(1))) +
    theme(axis.text.x= element_text(color = "blue4",size = rel(0.85)),
          axis.text.y = element_text(color = "blue4",size = rel(0.85))) +
    theme(legend.text = element_text(size = rel(0.8), color = "blue4"))}
```

# Introduction 

Energy policies are important for development of the country’s economy and production. With the growth of the economy and production, the requirement of the energy has increased in Turkey. As a reflection of the requirement, there has been an increase in energy production/consumption and a change in energy policies. The ratio of energy production to meet energy consumption in Turkey is low. According to 2017 data, the rate of imports in primary energy resources is 76 percent, means that Turkey is foreign-dependent in terms of energy (2020). This dependence on foreign energy is reflected in the current account balance negatively. Energy resources (fossil fuel) are insufficient in terms of energy production in Turkey, so renewable energy sources should be considered to improve current account balance.

Based on the situations that are mentioned, this project examines the relationship between energy production, consumption, and consumption between 1996 and 2020 with gross domestic product and current account balance. We expect a positive relationship between energy consumption and gross domestic product, and between the generation of renewable resources and gross domestic product. We expect a negative relationship between current account balance and energy consumption, as we are externally dependent on energy resources. We used World Bank Databese for the analysis.


# Analysis {.tabset}

## Sources of Electricity Production
```{r warning = FALSE,include=FALSE, echo =FALSE}
#data base
dtproduction1 <- data.table(
  WDI(country = c("TR"),
      indicator=c("EG.ELC.FOSL.ZS","EG.ELC.HYRO.ZS","EG.ELC.NUCL.ZS","EG.ELC.RNWX.ZS"),
      start = 1996,
      end= 2020,
      extra = FALSE)
)
clabsel <-   labs(title = "Electricity production from sources(% of total)",
                  subtitle = "Turkey",
                  x = "Years", y = "Sources",
                  caption = "Source: World Bank")

line1 <- ggplot(dtproduction1, mapping = aes(x = year)) + geom_line(size = 1, mapping = aes(y = EG.ELC.FOSL.ZS, color = "Fossil"))+
  geom_line(size = 1, mapping = aes(y = EG.ELC.NUCL.ZS, color = "nuclear")) +
  geom_line(size = 1, mapping = aes(y = EG.ELC.RNWX.ZS, color = "renewable"))+
  geom_line(size = 1, mapping = aes(y = EG.ELC.HYRO.ZS, color = "hydroelectric")) +
  guides(col = guide_legend("")) +clabsel + my_theme()

dtproduction <- data.table(
  WDI(country = c("TR"),
      indicator=c("EG.ELC.PETR.ZS","EG.ELC.NGAS.ZS","EG.ELC.COAL.ZS","EG.ELC.HYRO.ZS"),
      start = 1996,
      end= 2020,
      extra = FALSE)
)
clabsel <-   labs(title = "Electricity production from sources(% of total)",
                  subtitle = "Turkey",
                  x = "Years", y = "Sources",
                  caption = "Source: World Bank")

line2 <- ggplot(dtproduction, mapping = aes(x = year)) + 
  geom_line(size = 0.8, mapping = aes(y = EG.ELC.PETR.ZS, color = "Oil"))+
  geom_line(size = 0.8, mapping = aes(y = EG.ELC.NGAS.ZS, color = "Natural Gas")) +
  geom_line(size = 0.8, mapping = aes(y = EG.ELC.COAL.ZS, color = "Coal"))+
  geom_line(size = 0.8, mapping = aes(y = EG.ELC.HYRO.ZS, color = "Hydroelectric")) +
  guides(col = guide_legend("")) +clabsel + my_theme()


```

Energy consumption inevitably grows in parallel with the increasing population, urbanization, industrialization, widespread use of technology, and increased welfare. Hence, electric energy, which has become an indispensable need for human beings, has become a part of our daily life in many respects. Today to protect living standards, we need electrical energy in many sectors such as housing, industry, and transportation. The search for new energy sources for the generation of electricity has increased rapidly due to the economically inexpensive and environmentally friendly production of electrical energy. We can examine the methods commonly used to generate electrical power with current technologies in three main groups.

i) Fossil fuels (coal, oil, natural gas, etc.)

ii) Nuclear reactors

iii) Renewable energy sources

Despite using these three ways for electricity production in the world different from this situation in Turkey. As it can be seen below, when we look at electricity production from renewable sources (excluding hydroelectric), fossil sources (oil, gas, and coal), nuclear sources, and hydroelectric sources together, we observe that while the largest share is fossil fuels, nuclear and renewable sources have an incomparably small percentage in Turkey.

```{r, results=T, out.width="100%"}
ggplotly(line1)
```
                            Source: World Bank
                            
To investigate electricity energy production in Turkey, we will focus on the two most used sources are fossil and hydroelectric.  For this, we will get benefit from both the World Bank database and The Republic of Turkey's annual plans. While the first graph shows electricity production sources, the second one shows the frequency of using these in yearly plans.

```{r freqs}
clabs_f <-   labs(title = "Electricity production from sources(% of total) (Left) & Frequencies (Right)",
                 subtitle = "Turkey",
                 x = "Years", y = "Sources",
                 caption = "Source: Annual Plans")
gf <- ggplot(freqsDT, mapping = aes(x = Years)) +
  geom_point(size = 1, mapping = aes(y = petrol, color= "Oil"))+
  geom_point(size = 1, mapping = aes(y = gaz, color = "Natural Gas")) +
  geom_point(size = 1, mapping = aes(y = kömür, color = "Coal")) +
  geom_point(size = 1, mapping = aes(y = hidroelektrik, color = "Hydroelectric")) +
  my_theme() +
  scale_x_discrete(breaks=c(1995,2000,2005,2010,2015,2020))+
  guides(col = guide_legend("")) + clabs_f

```

```{r, results=T, figures-side, fig.show="hold", out.width="100%"}

p2<- subplot(
  add_trace(line2), 
  add_trace(gf),
  margin = 0.05)
p2
```

When we start analyzing the second graph from 1996, we observe that the use of oil word has an increasing trend, which means oil may have a decisive influence on energy policies from 1996 to 2002. It is not surprising that oil is discussed in energy policies during this period because although oil demand decreased due to the crisis in 1997 in Asian markets, OPEC increased the supply of oil, which ineluctably caused a decline in oil price. However, when we look at the first graph, we see that; The share of oil for electricity production in Turkey is relatively low and has a declining trend is moving in a straight line can be called stabile. The main reasons for the low share of oil for electricity generation, even though it is mentioned a lot in the annual program, are the volatility of oil prices and high production costs.


Apart from the oil source, which has an increasing slope, natural gas also strikes our eyes in the second graph. Natural gas has a special place and importance in the electricity sector. It is beneficial to continue the consumption development of natural gas, which is preferred due to its advantages in terms of price, efficiency, and environment, with a healthy understanding considering the principles of planning and resource diversity. And the first graph supports this. Between 1996 and 2001, we see that the source with the largest percentage is hydroelectricity, but it has a declining trend. On the other hand, natural gas has a rapid upward trend as an electricity generation source again for the same period. There are multiple reasons for this situation:

1.	Natural gas is essential in terms of security of supply as a backup solution to the problem of energy supply shortage because it can react quickly to electricity demand changes.

2.	Natural gas power plants have relatively low investment costs and fast installation times.

3.	In 1998, significant steps were taken in natural gas connections to meet the increasing energy demand.

Therefore, after 1999 we observe that natural gas has the most significant share in Turkey's electricity production. By 2019, we see that the share of coal and hydroelectricity in electricity generation has increased considerably.(2020)

When we continue with hydroelectric, which is the only remarkable renewable energy source used in Turkey, in the second graph hydroelectricity keeps its place in energy policies.However, we observe a sharp decline from 1996 to 2001 in the first graph. Then even if there are ups and downs in the following years, its share tends to decrease. The main reason behind this is related to being insufficient in investments. "Build-Operate-Transfer" and "Build-Operate" models were proposed to solve the problems arising in ensuring sustainability in energy investments. With those two models, it was aimed to find investors and encourage the private sector. However, this policy couldn’t reach the desired result. Even if hydroelectric power generation is the best source for Turkey, in recent years, the destruction of the forest ecosystem for establishing hydroelectric power plants has met with great public reaction.

Lastly, when we look at coal, we see very stable progress in both of the graphs. Even though there are ups and downs in the first graph, coal keeps its share around 30% between 1996 and 2015. There is a very similar case in the second graph; coal keeps its place in energy policies.

## Relationship Between Energy Consumption and GDP & GDP Growth

Energy is a necessary resource for countries, especially that have an increasing population, economic growth, urbanization, and technological developments. However, limited natural resources also concern these countries as the fossil fuels such as oil, coal and natural gas are limited and lead to a rise in energy prices. This urged countries to search for new energy resources and produce energy more efficiently from their resources. Keeping in mind the high dependency of developing countries on energy resources, the relationship between GDP and energy consumption levels became a relevant issue that is still debated. Even though the literature provides different results for the same countries because of data and different methodologies used in the studies in terms of definition and specification of variables, time span considered, econometric techniques employed, the economic growth and energy demand are linked. Turkey, as a developing country, relies on external resources for energy. According to the Ministry of Foreign Affairs of the Turkish Republic, Turkey has the highest rate of increasing energy demand among OECD countries since the mid-2000s. The researchers for McKinsey (2019) observed that when economies grow, they demand higher energy and the GDP growth reacts to a constraint in energy negatively in the country.

Throughout history, the economic changes in the energy sector and developments in the energy sector had impacted the policies regarding energy. Especially after the 1980s, the rapid increase in population and industrialization. Globalization has clearly created a higher need for energy consumption by prioritizing tourism and service sectors and thus created an additional demand for fossil fuels, natural gas and petroleum (Mucuk & Uysal, 2009). The intense levels of energy are required for factors of production to be used in increasing levels in production. The increasing energy demand causes an increase in more investment in energy and leads to an efficient use of energy and with the technological developments lead to a more productive labour market (Develi & Korkmaz, 2012). Therefore, energy is crucial for economic development for many countries including Turkey. Although the relationship between these two should be further investigated through regression and time series analysis, here the relation between GDP and energy consumption levels are observed through the last two decades and interpreted.

```{r}
gdpgr <-WDI(country = "TR", indicator = "NY.GDP.MKTP.KD.ZG", start = 1996, end = 2015 )
gdp <-WDI(country = "TR", indicator = "NY.GDP.MKTP.CD", start = 1996, end = 2015 )
energuse<- WDI(country = "TR", indicator = "EG.USE.PCAP.KG.OE", start = 1996, end = 2015 )
gdp$NY.GDP.MKTP.CD <- gdp$NY.GDP.MKTP.CD/1000000

head(gdpgr)
old1 <- c("iso2c", "country", "NY.GDP.MKTP.KD.ZG", "year")
old2 <- c("iso2c", "country", "NY.GDP.MKTP.CD", "year")
old3 <- c("iso2c", "country", "EG.USE.PCAP.KG.OE", "year")


setnames(gdpgr, old1, c("isocode", "country", "gdpgr", "year"))
setnames(gdp,old2, c("isocode", "country", "gdp", "year"))
setnames(energuse, old3, c("isocode", "country", "energyusee", "year"))

head(gdpgr)
head(gdp)
head(energuse)


dataa <- merge(energuse, gdpgr, by.x = "year", by.y = "year")
datab <- merge(energuse, gdp, by.x = "year", by.y ="year")

class(dataa)
head(dataa)
head(datab)


#Visualization--------- 
labss1 <- labs(title = "Relationship Between Enegry Consumption and GDP Growth",
               x = "Energy Consumption (kg of oil equivalent per capita)", y = "GDP growth (annual %)",
               caption = "Source: WorldBank")

labss2 <- labs(title = "Relationship Between Enegry Consumption and GDP",
               x = "Energy Consumption (kg of oil equivalent per capita)", y = "GDP in million (current US$)",
               caption = "Source: WorldBank")

linediag1 <- ggplot(data = datab, aes(x = energyusee, y= gdp, color= year)) +geom_point() +my_theme() +labss2 + 
  scale_y_continuous(labels = comma) +geom_smooth(method="lm", se=F, color="purple")

linediag2 <- ggplot(data = dataa, aes(x = energyusee, y= gdpgr, color= year)) +geom_point() +my_theme() +labss1 + 
  scale_y_continuous() +geom_smooth(method="lm", se=F, color="purple")

dtgdpc <- 
  WDI(country = c("TR"),
      indicator="EG.GDP.PUSE.KO.PP.KD",
      start = 1996,
      end= 2015,
      extra = FALSE)

head(dtgdpc)

old4 <- c("iso2c", "country", "EG.GDP.PUSE.KO.PP.KD", "year")
setnames(dtgdpc, old4, c("isocode", "country", "gdpperenergy", "year"))

labss3 <- labs(title = "GDP per Energy Consumed in Turkey",
               x = "Years", y = "GDP per Energy Consumed",
               caption = "Source: WorldBank")
head(dtgdpc)

linediag3 <- ggplot(data = dtgdpc, aes(x = year, y= gdpperenergy)) +geom_line(color="purple") +my_theme() +labss3
```

```{r, results=T, out.width="100%"}
ggplotly(linediag1)
```

The first graph shows the relationship between GDP and energy consumption levels in Turkey. The positive slope of the trend line indicates the high correlation between these two variables. As the years go by we observe an even stronger bond between GDP and energy consumption levels as the recent years are allocated near to the higher part of the line. There are some exceptions to this situation which are mostly the years that an economic crisis occurred. These years are allocated at lower parts of the trend line. These were indeed expected because Turkey and its economy rely highly on energy consumption. As it was mentioned in the literature review part, energy consumption is necessary for efficient use of the resources, boosts labor productivity and hence leads to a higher total level of goods produced within Turkey. This relationship is even stronger as Turkey is a developing country trying to improve at industrialization, urbanization and globalization aspects. 

```{r, results=T, out.width="100%"}
ggplotly(linediag2)
```

Similar to the results above, the relationship between the energy consumption levels and GDP growth is positive as we observe the trend line in this graph has a positive slope. However, the positive relation is more gradual in this case because GDP growth is not as sensitively affected by the energy consumed. Most of the years fall within a small range of the trend but there are some years that are far away from the rest and look like outliers. The years that are far under the trend are the years where major economic crises occurred and the 2 years followed by them. We see the second year after the economic crises at the lowest part as these are the years where economic growth was still negative but energy consumption levels were extraordinarily high to get the economy out of the crises. Similarly, the years after the crises occurred to take place in the lower part of the graph than the years that crises started. The remaining years that are above and very close to the trend line can be used to prove the close relationship between GDP growth and energy consumption levels in Turkey, although their relationship is not as sensitive as the relation between GDP and energy consumption levels. 

```{r, results=T, out.width="100%"}
ggplotly(linediag3)
```

This analysis can be further conducted by looking at the proportional change of GDP and energy usage through the years. By dividing the GDP levels with energy use for each year, we observe the relation of how efficiently energy use contributed to GDP levels. The graph has an overall increasing trend and has fluctuations in 2000, 2005 and 2009. The economic crises in these years affected the increasing trend of these negatively as the GDP has lowered due to economic crises but the energy consumption levels remained at similar levels. It can be observed that the rate of increase is lower for the years right after the crises and is higher after generally after 2 years’ period. These two years are the years that the economy needs to get back to its track again after the crises occurred and are able to grow at it is potential levels again. 

## Relationship Between Energy Consumption and Current Account Balance & Manufacturing
```{r final_graphs}
aa <-WDIsearch("current account")
bb <-WDIsearch("manufacturing")
cc <-WDIsearch("energy")

cadTR <- WDI(country = "TR", indicator = "BN.CAB.XOKA.CD", start = 1996, end = 2020 )
#Current account balance (BoP)

manufTR <-WDI(country = "TR", indicator = "NV.IND.MANF.CN", start = 1996, end = 2020 )
#Manufacturing, value added (current LCU)

energconsTR <- WDI(country = "TR", indicator = "1.1_TOTAL.FINAL.ENERGY.CONSUM", start = 1996, end = 2020 )
#Total final energy consumption (TFEC)

old1 <- c("iso2c", "country", "BN.CAB.XOKA.CD", "year")
old2 <- c("iso2c", "country", "NV.IND.MANF.CN", "year")
old3 <- c("iso2c", "country", "1.1_TOTAL.FINAL.ENERGY.CONSUM", "year")

setnames(cadTR,old1, c("isocode", "country", "cab", "year"))
setnames(manufTR, old2, c("isocode", "country", "manuf", "year"))
setnames(energconsTR, old3, c("isocode", "country", "energy", "year"))

energconsTR$energy <- round(energconsTR$energy, digits = 3)
cadTR$cab <- cadTR$cab/100000

dt1 <- merge(cadTR, energconsTR, by.x = "year", by.y = "year")

dt2 <- merge(manufTR, energconsTR, by.x = "year", by.y = "year")

#Visualization--------- 

labss1 <- labs(title = "Relationship Between Current Account Balance and Energy Consumption",
              x = "Energy Consumption", y = "Current Account Balance (in million)",
              caption = "Source: WorldBank")

labss2 <- labs(title = "Relationship Between Manufacturing and Energy Consumption",
               x = "Energy Consumption", y = "Manufacturing",
               caption = "Source: WorldBank")

bardiag <- ggplot(data = dt1, aes(x=energy, y = cab , color= year )) +geom_point()  +labss1 +my_theme() +
  geom_smooth(method = "lm", se=F, color="purple") +scale_y_continuous(labels = comma)

bardiagg <- ggplot(data = dt2, aes(x=energy, y = manuf , color= year )) +geom_point()  +labss2 +my_theme() +
  geom_smooth(method = "lm", se=F, color="purple") +scale_y_continuous(labels = comma)
```

We know from other researches that, while the increase in electricity consumption causes economic growth, the increase in energy consumption based on imports used in electricity production affects the CAB and causes a current account deficit. Also, we know that energy use of industrial (manufacturing) sector has highest share in electric consumption of Turkey. (Oğulata,2002) Therefore, in this part we examine both the relationship between current account balance and energy consumption, and the relationship between manufacturing and energy consumption.

```{r results=TRUE, fig.width=7, fig.height=5, fig.align="right"}
ggplotly(bardiagg)
```

The graph shows that there is a positive relationship between manufacturing and energy consumption. This is an expected result that is not so surprising. Because of the energy consumption of industry has a highest share, while the manufacturing activities increase, electric consumption also increases. 

```{r results=TRUE, fig.width=9, fig.height=6}
ggplotly(bardiag)
```

The graph shows that there is a negative relationship between current account balance (CAB) and energy consumption. This is an also an expected result, because the ratio of energy production to meet energy consumption in the Turkey is low. The rate of imports in primary energy resources is 76 percent, means that Turkey is foreign-dependent in terms of energy Therefore, our energy sector depends on imports, which deteriotes CAB of Turkey. (2020)

To examine the energy supply of Turkey in detail, the resources to use or generate electricity are supplied from abroad. For example, natural gas is second place in primary energy supply, and we import 99% of the natural gas used in our country. (2020) To decrease the foreign-dependency, National Energy and Mining Policy and Renewable Energy Investments Encouragement Policy have been applied since 2019. (2020 Annual Plan) Exploration studies for natural gas, which is an important step in reducing foreign dependency, continue. Also, Nuclear Power Plant has been considered and works have started as of 2008.It would be better to consider renewable energy policies for the healthy world.

When these results are taken into account, it has been revealed that electricity should be produced by using national energy resources in order to ensure economic growth and to increase the amount of electricity consumption and prevent the current account deficit. (Gökten and Karatepe, 2016) Turkey is in a deficient situation in terms of resources, so to decrease the current account deficit and to ensure a stable economic growth renewable energy sources could be used such as wave, wind, solar and so on. Green energy and other energy generation facilities should be considered.  

## Relationship Between Renewable Energy and GDP per Capita

```{r renewable}
elecTR <-WDI(country = "TR", indicator = "4.1.2_REN.ELECTRICITY.OUTPUT", start = 1996, end = 2020 )

renewTR <- WDI(country = "TR", indicator = "NY.GDP.PCAP.KD", start = 1996, end = 2020)

old1 <- c("iso2c", "country", "4.1.2_REN.ELECTRICITY.OUTPUT", "year")
old2 <- c("iso2c", "country", "NY.GDP.PCAP.KD", "year")


setnames(elecTR,old1, c("isocode", "country", "electric", "year"))
setnames(renewTR, old2, c("isocode", "country", "renewable", "year"))


dataa <- merge(elecTR, renewTR, by.x = "year", by.y = "year")
```

```{r renewable graph }
labss1 <- labs(title = "Relationship Between Renewable Enegry and GDP per Capita",
               x = "Renewable energy electricity output (GWh)", y = "GDP per capita (constant 2010 US$)",
               caption = "Source: WorldBank")


bardiag <- ggplot(data = dataa, aes(x= electric, y = renewable, color= year )) +geom_point() +my_theme() +
  geom_smooth(method = "lm",se=F, color="purple") +scale_y_continuous(labels = comma) +labss1

```

```{r, results=T,  fig.width=9, fig.height=6}
ggplotly(bardiag)
```


Significant developments have been seen in renewable energy on a global scale, especially in the last 15 years. Turkey wants to improve the high rate of economic growth of recent years which is dependent on foreign supplies around 70% of primary energy sources. In other words, addiction is a risk to economic independence from time to time. For this reason, new and alternative energy sources are needed to meet the increasing energy needs, (Coban and Yorgancila,2018).  Around the world, countries have started to encourage using renewable energy sources and to prevent using of fossil fuels by increasing the cost of using them. Hence, the share of renewable energy in total supply has been increased in the last years. The rate of increase was 56 percent for OECD member states and more than 86 percent for high-income European countries. (Saygılı,2018) Even though the rate of using renewable energy is far from the optimal level, Turkey has a higher rate than the OECD average. 

As of 2020, Turkey's renewable energy installed capacity is 52.429 MW. While other countries generally use a diversified energy mix, Turkey mostly produces hydroelectric power. The highest share among the renewable energy sources belongs to hydraulic energy with 60.4%, which means 31.688 MW. This is followed by wind with 9.633 MW, sun with 7.750MW, and others with 3.358 MW. Given these data, the share of renewable energy in the total energy consumption of Turkey has amounted to 34.39%. Although the sharing of renewable energy sources has been increasing by every year, Turkey has struggled to invest enough in geographically abundant energy, such as solar and wind power, relative to Italy and Greece. Using of wind energy has been rising since 2009 but there is still a long way to go. Solar energy has only been around since 2014. These resources can generate electricity at a much lower cost than other sources, (Akarsu, Korucu,2019).

Moreover, as we can see at graph, while GDP per capita is increasing, using of renewable energy is also increasing. However, this relation is not a causal relation because in developing countries, renewable energy may not be priority for them, when income increases; they may prefer to make investment related to health, education, and other important areas rather than energy sector. Moreover, even they prefer to make energy investment, for low-income countries, economic growth may be much more important compared to environmental sustainability, as fossil fuels are much more cost competitive compared to renewables, fossil fuels can be used to promote economic growth (Saygılı,2018).


# Conclusion

Energy policy is one of the most fundamental public policies of countries because it affects countries in environmental, economic, and international relations. Therefore, in this study, we focused on the energy policies set in line with environmentally friendly energy produced, ensuring the security of energy supply, maintaining price stability, and examining their relations with economic parameters in Turkey. For this purpose, we have discussed energy policies under four main headings. 

In the first part, we have focused on sources of electricity production. By, doing this we have reached that Turkey has used mostly fossil sources and hydroelectricity to generate electricity. When we got a little deeper, we realized that natural gas is the primary energy production source because it is more environmentally friendly than other fossil resources. Its installation cost is low, and it can react to rapid changes in energy demand. In the second part, we have focused on the relationship between economic growth and energy use in Turkey. A close relationship between energy consumption levels with GDP and GDP growth for Turkey is observed through years as population, industrialization, productivity due to accelarated technological developments and capital investments. We conclude that there is a two-way relationship between energy use and economic growth. While the increase in economic growth triggers energy consumption, increasing energy consumption contributes to the economy's growth. Between 1996 and 2020, both energy consumption and GDP have increased parallelly over time in Turkey. 

In the third part, we have focused on the relationship between energy consumption and Current Account Balance & Manufacturing. By investigating these two relations, we aimed to support our claim in the second part and see how a country dependent on imported energy affects the current account balance. For this part, we conclude that while there is a positive relationship between manufacturing and energy consumption, our research shows a negative relationship between current account balance (CAB) and energy consumption. This leads us to the following point: Turkey is needed to give the current account deficit to continue to grow in production due to energy import. In the last part, we have focused on the renewable energy concept because we have thought that awareness of renewable energy has started to increase with the changing social structure. Even if not used much in Turkey, we are aware that the world's future depends on using renewable resources. In addition to this, there is a positive relationship between the economic development of a country and renewable energy production. For this part, we have concluded that we observed a positive relationship between those two variables. Turkey has increased the use of renewable energy sources in time. The main reason behind that, with the increase in economic development, the use of renewable energy increases. If we get everything together, with this research, we reach that Turkey’s energy policies between 1996 and 2020 are compatible with the country's economic movements.

# References

Çoban A, O. Y. (n.d.). Relationship between Renewable Energy Consumption and Sustainable Economic Growth: The Case of Turkey. .

H, S. (n.d.). RENEWABLE ENERGY USE IN TURKEY . Retrieved from https://www.tcmb.gov.tr/wps/wcm/connect/blog/en/main+menu/analyses/renewable_energy_us_+in_turkey

Karatepe, S. G. (2016). Electricity consumption and economic growth:A causality analysis for Turkey in the frame of import-based energy consumption and current account deficit. Energy Sources, Part B: Economics, Planning, and Policy, 385-389.

Lise K., W. M. (2007). Energy consumption and GDP in Turkey: Is there a co‐integration relationship?. Energy Economics.

Akarsu N.,Korucu G.(2019). What are the Main Determinants of Renewable Energy Consumption? A Panel Threshold Regression Approach. . Anadolu Üniversitesi Sosyal Bilimler Dergisi.

Oğulata, T. (2002). Sectoral energy consumption in Turkey. Renewable and Sustainable Energy Reviews.

Strateji ve Bütçe Başkanlığı. (2020). 1996-2020 Yıllık Programları. Retrieved from https://www.sbb.gov.tr/yillik-programlar/

UNION OF CHAMBERS OF TURKISH ENGINEERS AND ARCHITECTS (UCTAE) CHAMBER OF MECHANICAL ENGINEERS (CME) ENERGY COMMISSION. (2020). Turkey Energy Outlook. 

WorldBank. Data of Turkey.

Develi M.,Korkmaz Ö. (2012). Türkiye'de Birincil Enerji Kullanımı, Üretimi ve Gayri Safi Yurt İçi Hasıla (GSYİH) Arasındaki İlişki. Dokuz Eylül Üniversitesi İktisadi İdari Bilimler Fakültesi Dergisi, 1-25.

Sharma N., Smeets B., Tryggestad C. (2019). The Decoupling of GDP and Energy Growth: A CEO Guide. McKinsey Quarterly .

Mucuk M., Uysal D. (2009). Türkiye Ekonomisinde Enerji Tüketimi ve Ekonomik Büyüme. Maliye Dergisi , 105-115.
