---
title: "grafikler toplu"
author: "İrem Kaya"
date: "02 02 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, results = F)
```

```{r run_scripts}
source("R/01_download_pdf.R", encoding = "UTF-8")
source("R/02_read_pdf.R")
```

```{r search_keywords}

search_terms <- c("yenilenebilir", "rüzgar", "rüzgâr", "güneş", "nükleer",
                  "hidroelektrik", "jeotermal", "ngs", "hes", "petrol", "gaz", 
                  "kömür", "doğalgaz", "linyit")

freqs <- findFreq(search_terms)

freqsDT <- as.data.table(freqs, keep.rownames = "Years")
freqsDT$gaz <- freqsDT$gaz + freqsDT$doğalgaz
freqsDT$hidroelektrik <- freqsDT$hidroelektrik + freqsDT$hes
freqsDT$nükleer <- freqsDT$nükleer + freqsDT$ngs
freqsDT$kömür <- freqsDT$kömür + freqsDT$linyit
# sums[search_terms]

# findFreq("enerji")
# sums["enerji"]

# # all words starting with verim Adjust regex to find what you need.
# freq[,grep("^verim", colnames(freq))]
# sums[grep("^verim", names(sums))]
```

```{r libraries}
my_packages <- c("WDI", "data.table", "ggplot2", "plotly", "CBRT", "ggthemes", "scales")
not_installed <- my_packages[!(my_packages %in% installed.packages()[ , "Package"])]
if(length(not_installed)) install.packages(not_installed, repos = "http://cran.us.r-project.org")

library(WDI)
library(data.table)
library(ggplot2)
library(plotly)
library(CBRT)
library(ggthemes)
library(scales)
```

```{r theme}
# 
# my_theme <- function(base_size = 11,base_family ="",cols =c("#EB5E57","#9D3BED"))
# {theme(rect= element_blank(),
#        text=element_text(size=11,family = "latha"),
#        plot.title= element_text(size =rel(1.4), color= "#FA6850",face = "bold"),
#        plot.subtitle = element_text(size =rel(1), color= "#FA6850", face= "bold"),
#        plot.caption = element_text(color = "#FA6850", hjust = 1,vjust = 1,size = rel(0.7)),
#        panel.background = element_rect(colour = "#E9C2F0"),
#        panel.grid.major.y = element_line( color = "#A462FA",linetype = "dotted", size =rel(0.3)),
#        panel.grid.major.x = element_blank(),
#        panel.grid.minor.y = element_blank(),
#        panel.grid.minor.x = element_blank(),
#        panel.border = element_blank(),
#        axis.title.x = element_text(colour= "#910EF5",size = rel(1)),
#        axis.title.y = element_text(colour= "#910EF5",size = rel(1)),
#        axis.line.x = element_line(colour = "#FA6850", size = rel(0.6)),
#        axis.line.y = element_line(colour = "#FA6850",size = rel(0.6)))}


my_theme <- function()  { theme_bw()   +
    theme(panel.border = element_rect(color = "gray70", size = 1.5)) +
  theme(legend.position = "none") +
  theme(plot.caption = element_text(color = "blue4", size = rel(0.9))) +
  theme(panel.background = element_rect(fill = "mintcream", colour = "gray70"), panel.grid = element_line(colour ="gray70"),
      panel.grid.minor = element_line(size = rel(0.5)),
      strip.background = element_rect(fill = "mintcream",
                              colour = "gray70"), complete = TRUE) +theme(axis.title.x= element_text(color = "blue4", size = rel(0.9)),
                            axis.title.y = element_text(color = "blue4", size = rel(0.9))) +
    theme(plot.title = element_text(color = "blue4", size = rel(1))) +
    theme(axis.text.x= element_text(color = "blue4",size = rel(0.85)),
          axis.text.y = element_text(color = "blue4",size = rel(0.85))) +
    theme(legend.text = element_text(size = rel(0.8), color = "blue4"))}
```

## Analysis {.tabset}

### name : oil rents (% of GDP)
oil rents (Oil rents are the difference between the value of crude oil production at world prices and total costs of production.)
```{r oil_rents}
WDIsearch(string = "Oil rents",
          field = "name",
          short = FALSE)

df <- data.table(
  WDI(country = c("TR"),
      indicator="NY.GDP.PETR.RT.ZS",
      start = 1996,
      end= 2020,
      extra = FALSE)
)
head(df)
tail(df)

clabs <-   labs(title = "Oil Rents (% of GDP)",
                subtitle = "Turkey",
                x = "Years", y = "Oil Rents",
                caption = "Source: World Bank")

g1 <- ggplot(df, aes(year, NY.GDP.PETR.RT.ZS)) + geom_line() +scale_y_log10()+ my_theme() + clabs
g2 <- ggplot(df, aes(year, NY.GDP.PETR.RT.ZS)) + geom_line() + my_theme() + clabs

```

```{r, results=T}
ggplotly(g1)
ggplotly(g2)
```

### rents
(oil, natural gas, total natural resource) tüm rentleri bir araya topladım
```{r rents}
WDIsearch(string = "rents",
          field = "name",
          short = FALSE)

dfrents <- data.table(
  WDI(country = c("TR"),
      indicator=c("NY.GDP.PETR.RT.ZS","NY.GDP.COAL.RT.ZS","NY.GDP.NGAS.RT.ZS","NY.GDP.TOTL.RT.ZS"),
      start = 1996,
      end= 2020,
      extra = FALSE)
)

clabs7 <-   labs(title = "Rents(% of GDP)",
                 subtitle = "Turkey",
                 x = "Years", y = "Sources",
                 caption = "Source: World Bank")
g3 <- ggplot(dfrents, mapping = aes(x = year)) + geom_line(size = 1, mapping = aes(y = NY.GDP.PETR.RT.ZS, color = "Oil"))+
  geom_line(size = 1, mapping = aes(y = NY.GDP.COAL.RT.ZS, color = "Coal")) +
  geom_line(size = 1, mapping = aes(y = NY.GDP.NGAS.RT.ZS, color = "Natural Gas"))+
  geom_line(size = 1, mapping = aes(y = NY.GDP.TOTL.RT.ZS, color = "total natural resources")) +my_theme() +
  guides(col = guide_legend("")) +clabs7
```

```{r, results=T}
ggplotly(g3)
```

```{r freqs, results=T}
clabs_f <-   labs(title = "Frequencies",
                 subtitle = "Turkey",
                 x = "Years", y = "Sources",
                 caption = "Source: Annual Plans")
gf <- ggplot(freqsDT, mapping = aes(x = Years)) +
  geom_point(size = 1, mapping = aes(y = petrol, color= "petrol"))+
  geom_point(size = 1, mapping = aes(y = kömür, color = "kömür")) +
  geom_point(size = 1, mapping = aes(y = gaz, color = "gaz")) +
  geom_point(size = 1, mapping = aes(y = hidroelektrik, color = "hidroelektrik")) +
  my_theme() +
  guides(col = guide_legend("")) + clabs_f

# gf <- ggplot(freqsDT, aes(x = Years, y = value, colour = variable)) + 
#     geom_smooth() + geom_point() + 
#     my_theme() +
#   guides(col = guide_legend("")) + clabs_f

# gf <- ggplot(freqsDT, mapping = aes(x = Years)) +
#   geom_line(size = 1, mapping = aes(y = kömür), color = "black")+
#   geom_line(size = 1, mapping = aes(y = petrol), colour = "dark green") +
#   geom_line(size = 1, mapping = aes(y = gaz), colour = "blue")+
#   my_theme() +
#   guides(col = guide_legend("")) +clabs_f

ggplotly(gf)
```

###  alternative and nuclear energy (% of total energy use)
```{r alternative_nuclear}
WDIsearch(string = "Alternative and nuclear energy",
          field = "name",
          short = FALSE)

dtnuclear <- data.table(
  WDI(country = c("TR"),
      indicator="EG.USE.COMM.CL.ZS",
      start = 1996,
      end= 2020,
      extra = FALSE)
)
 head(dtnuclear)
clabs1 <-   labs(title = "Alternative and Nuclear Energy (% of total energy use)",
                subtitle = "Turkey",
                x = "Years", y = "Alt. and Nuc. Energy",
                caption = "Source: World Bank")


g4 <- ggplot(dtnuclear, aes(year, EG.USE.COMM.CL.ZS)) + geom_line() +scale_y_log10()+ my_theme() + clabs1
g5 <- ggplot(dtnuclear, aes(year, EG.USE.COMM.CL.ZS)) + geom_line() + my_theme() + clabs1
```

```{r, results=T}
ggplotly(g4)
ggplotly(g5)
```

### bu noktada karşılaştırma yapalım (middle esat and north africa)
```{r compare_ME_NA}
dtnuclear2 <- data.table(
  WDI(country = c("MEA"),
      indicator="EG.USE.COMM.CL.ZS",
      start = 1996,
      end= 2020,
      extra = FALSE)
)

DTnuc2 <- merge(dtnuclear, dtnuclear2, by = "year", all.x = TRUE, all.y = T)
head(DTnuc2)

clabsnuc <-   labs(title = "Alternative and Nuclear Energy (% of total energy use)",
                 subtitle = "Turkey and Middle East and North Africa",
                 x = "Years", y = "Alt. and Nuc. Energy(%)",
                 caption = "Source: World Bank")

g6 <- ggplot(DTnuc2, mapping = aes(x = year)) + geom_bar(aes(y = EG.USE.COMM.CL.ZS.x, color= "Turkey"), stat ="identity",position =position_dodge())+
  geom_line(size = 1, mapping = aes(y = EG.USE.COMM.CL.ZS.y, color = "Middle East & North Africa")) + my_theme() +
  guides(col = guide_legend(""))+ clabsnuc
```

```{r, results=T}
ggplotly(g6)
```

### bir de middle income grupla karşılaştıralım
```{r compare_MI}
dtnuclear3 <- data.table(
  WDI(country = c("MIC"),
      indicator="EG.USE.COMM.CL.ZS",
      start = 1996,
      end= 2020,
      extra = FALSE)
)

DTnuc3 <- merge(dtnuclear, dtnuclear3, by = "year", all.x = TRUE, all.y = T)
head(DTnuc3)

clabsnuc3 <-   labs(title = "Alternative and Nuclear Energy (% of total energy use)",
                   subtitle = "Turkey and Middle Income",
                   x = "Years", y = "Alt. and Nuc. Energy",
                   caption = "Source: World Bank")

head(DTnuc3)
g7 <- ggplot(DTnuc3, mapping = aes(x = year)) + 
  geom_line(size = 1, mapping = aes(y = EG.USE.COMM.CL.ZS.x, color = "Turkey")) +
  geom_line(size = 1, mapping = aes(y = EG.USE.COMM.CL.ZS.y, color = "Middle Income")) + 
  my_theme() +
  guides(col = guide_legend("")) + 
  clabsnuc3

g8 <- ggplot(DTnuc3, mapping = aes(x = year)) + 
  geom_bar(aes(y = EG.USE.COMM.CL.ZS.x, color ="Turkey"), stat ="identity",position =position_dodge()) +
  geom_line(size = 1, mapping = aes(y = EG.USE.COMM.CL.ZS.y, color = "Middle Income")) + 
  my_theme() +
  guides(col = guide_legend(""))+ 
  clabsnuc3
```

```{r, results=T}
ggplotly(g7)
ggplotly(g8)
```

###  energy imports,net(% of energy use)
```{r energy_imports}
WDIsearch(string = "energy imports",
          field = "name",
          short = FALSE)

dtimport <- data.table(
  WDI(country = c("TR"),
      indicator="EG.IMP.CONS.ZS",
      start = 1996,
      end= 2020,
      extra = FALSE)
)
head(dtimport)
clabso <-   labs(title = "Energy Imports,net(% of energy use)",
                 subtitle = "Turkey and OECD",
                 x = "Years", y = "Enegy Import",
                 caption = "Source: World Bank")
dtimportoecd <- data.table(
  WDI(country = c("OED"),
      indicator="EG.IMP.CONS.ZS",
      start = 1996,
      end= 2020,
      extra = FALSE))
DTimport <- merge(dtimport, dtimportoecd, by = "year", all.x = TRUE, all.y = T)
head(DTimport)

g9 <- ggplot(DTimport, mapping = aes(x = year)) + geom_line(size = 1, mapping = aes(y = EG.IMP.CONS.ZS.x, color = "Turkey"))+
  geom_line(size = 1, mapping = aes(y = EG.IMP.CONS.ZS.y, color = "OECD")) + my_theme() +
  guides(col = guide_legend(""))+ clabso

g10 <- ggplot(DTimport, mapping = aes(x = year)) + geom_bar(aes(y = EG.IMP.CONS.ZS.x, color= "Turkey"), stat ="identity",position =position_dodge())+
  geom_line(size = 1, mapping = aes(y = EG.IMP.CONS.ZS.y, color = "OECD")) + my_theme() +
  guides(col = guide_legend(""))+ clabso
```

```{r, results=T}
ggplotly(g9)
ggplotly(g10)
```
### middle income ile karşılaştıralım
```{r compare_mi2}

dtimportmid <- data.table(
  WDI(country = c("MIC"),
      indicator="EG.IMP.CONS.ZS",
      start = 1996,
      end= 2020,
      extra = FALSE))
DTimport2 <- merge(dtimport, dtimportmid, by = "year", all.x = TRUE, all.y = T)
head(DTimport2)
clabsx <-   labs(title = "Energy Imports,net(% of energy use)",
                 subtitle = "Turkey and Middle Income",
                 x = "Years", y = "Enegy Import",
                 caption = "Source: World Bank")
g11 <- ggplot(DTimport2, mapping = aes(x = year)) + geom_line(size = 1, mapping = aes(y = EG.IMP.CONS.ZS.x, color = "Turkey"))+
  geom_line(size = 1, mapping = aes(y = EG.IMP.CONS.ZS.y, color = "Middle Income")) + my_theme() +
  guides(col = guide_legend(""))+ clabsx
ggplotly(g11)
```

```{r, results=T}
ggplotly(g11)
```

###  bir de dünya ile karşılaştıralım
```{r compare__world}
dtimportworld <- data.table(
  WDI(country = c("WLD"),
      indicator="EG.IMP.CONS.ZS",
      start = 1996,
      end= 2020,
      extra = FALSE))
DTimport3 <- merge(dtimport, dtimportworld, by = "year", all.x = TRUE, all.y = T)
head(DTimport3)

clabs3 <-   labs(title = "Energy Imports,net(% of energy use)",
                 subtitle = "Turkey and World",
                 x = "Years", y = "Enegy Import",
                 caption = "Source: World Bank")


g12 <- ggplot(DTimport3, mapping = aes(x = year)) + geom_line(size = 1, mapping = aes(y = EG.IMP.CONS.ZS.x, color = "Turkey"))+
  geom_line(size = 1, mapping = aes(y = EG.IMP.CONS.ZS.y, color = "World")) + my_theme() +
  guides(col = guide_legend(""))+ clabs3

g13 <- ggplot(DTimport3, mapping = aes(x = year)) + geom_bar(aes(y = EG.IMP.CONS.ZS.x, color="Turkey"), stat ="identity",position =position_dodge())+
  geom_line(size = 1, mapping = aes(y = EG.IMP.CONS.ZS.y, color = "World")) + my_theme() +
  guides(col = guide_legend(""))+ clabs3
```

```{r, results=T}
ggplotly(g12)
ggplotly(g13)
```

### electricity production from... (% of total)
```{r electricit_prod}
WDIsearch(string = "electricity production from",
          field = "name",
          short = FALSE)

dtproduction <- data.table(
  WDI(country = c("TR"),
      indicator=c("EG.ELC.PETR.ZS","EG.ELC.NGAS.ZS","EG.ELC.COAL.ZS","EG.ELC.HYRO.ZS"),
      start = 1996,
      end= 2020,
      extra = FALSE)
)
clabsel <-   labs(title = "Electricity production from sources(% of total)",
                 subtitle = "Turkey",
                 x = "Years", y = "Sources",
                 caption = "Source: World Bank")

g14 <- ggplot(dtproduction, mapping = aes(x = year)) + geom_line(size = 1, mapping = aes(y = EG.ELC.PETR.ZS, color = "Oil"))+
  geom_line(size = 1, mapping = aes(y = EG.ELC.NGAS.ZS, color = "natural gas")) +
  geom_line(size = 1, mapping = aes(y = EG.ELC.COAL.ZS, color = "Coal"))+
  geom_line(size = 1, mapping = aes(y = EG.ELC.HYRO.ZS, color = "hydroelectric")) +my_theme() +
  guides(col = guide_legend("")) +clabsel
```

```{r, results=T}
ggplotly(g14)
```

### fossil fuel energy consumption (% of total)
```{r fossil_fuel}
WDIsearch(string = "electricity production from",
          field = "name",
          short = FALSE)

dtfossil1 <- data.table(
  WDI(country = c("TR"),
      indicator="EG.IMP.CONS.ZS",
      start = 1996,
      end= 2020,
      extra = FALSE))

dtfossil2 <- data.table(
  WDI(country = c("MIC"),
      indicator="EG.IMP.CONS.ZS",
      start = 1996,
      end= 2020,
      extra = FALSE))

DTfossil <- merge(dtfossil1, dtfossil2, by = "year", all.x = TRUE, all.y = T)

clabs4 <-   labs(title = "Fossil Fuel Energy Consumption (% of total)",
                 subtitle = "Turkey and Middle Income Countries",
                 x = "Years", y = "Condumption",
                 caption = "Source: World Bank")

head(DTfossil)
g15 <- ggplot(DTfossil, mapping = aes(x = year)) + geom_line(size = 1, mapping = aes(y = EG.IMP.CONS.ZS.x, color = "Turkey"))+
  geom_line(size = 1, mapping = aes(y = EG.IMP.CONS.ZS.y, color = "Middle Income")) + my_theme() +
  guides(col = guide_legend(""))+ clabs4

g16 <- ggplot(DTfossil, mapping = aes(x = year)) + geom_bar(aes(y = EG.IMP.CONS.ZS.x, color ="Turkey"), stat ="identity",position =position_dodge())+
  geom_line(size = 1, mapping = aes(y = EG.IMP.CONS.ZS.y, color = "OECD")) + my_theme() +
  guides(col = guide_legend(""))+ clabs4
```

```{r, results=T}
ggplotly(g15)
ggplotly(g16)
```

### bir de Europe & Central Asia ile karşılaştıralım
```{r compare_europe_ca}
dtfossil3 <- data.table(
  WDI(country = c("ECS"),
      indicator="EG.IMP.CONS.ZS",
      start = 1996,
      end= 2020,
      extra = FALSE))

DTfossil2 <- merge(dtfossil1, dtfossil3, by = "year", all.x = TRUE, all.y = T)

clabs5 <-   labs(title = "Fossil Fuel Energy Consumption (% of total)",
                 subtitle = "Turkey and Europe & Central Asia",
                 x = "Years", y = "Consumption",
                 caption = "Source: World Bank")

head(DTfossil2)
g17 <- ggplot(DTfossil2, mapping = aes(x = year)) + geom_line(size = 1, mapping = aes(y = EG.IMP.CONS.ZS.x, color = "Turkey"))+
  geom_line(size = 1, mapping = aes(y = EG.IMP.CONS.ZS.y, color = "Europe & Central Asia")) + my_theme() +
  guides(col = guide_legend(""))+ clabs5
```

```{r, results=T}
ggplotly(g17)
```

###  tek başına türkiye daha anlamlı
```{r turkey}
g18 <- ggplot(DTfossil2, mapping = aes(x = year)) + geom_line(size = 1, mapping = aes(y = EG.IMP.CONS.ZS.x, color = "Turkey"))+
  my_theme()+ clabs5
```

```{r, results=T}
ggplotly(g18)
```

###  combustible renewable and waste (% of total)
```{r combustible_renewable_and_waste}
WDIsearch(string = "Combustible renewable",
          field = "name",
          short = FALSE)

dtwaste <- data.table(
  WDI(country = c("TR"),
      indicator="EG.USE.CRNW.ZS",
      start = 1996,
      end= 2020,
      extra = FALSE)
)
head(dtwaste)


clabs <-   labs(title = "Combustible Renewable and Waste (% of total)",
                subtitle = "Turkey",
                x = "Years", y = "",
                caption = "Source: World Bank")

g19 <- ggplot(dtwaste, aes(year, EG.USE.CRNW.ZS)) + geom_line() + my_theme() + clabs
```

```{r, results=T}
ggplotly(g19)
```

### gdp growth ve yenilenebilir enerji tüketimi (%)
```{r renewable_energy_consumption}
WDIsearch(string = "renewable energy",
          field = "name",
          short = FALSE)

dfrenew <- data.table(
  WDI(country = c("TR"),
      indicator="2.1_SHARE.TOTAL.RE.IN.TFEC",
      start = 1996,
      end= 2020,
      extra = FALSE)
)
head(dfrenew)

```

```{r gdp_growth}
WDIsearch(string = "GDP growth",
          field = "name",
          short = FALSE)

dtgdp <- data.table(
  WDI(country = c("TR"),
      indicator="NY.GDP.MKTP.KD.ZG",
      start = 1996,
      end= 2020,
      extra = FALSE)
)

head(dfrenew)

setnames(dfrenew, c("2.1_SHARE.TOTAL.RE.IN.TFEC"), c("renew"))

head(dfrenew)

head(dtgdp)
setnames(dtgdp, c("NY.GDP.MKTP.KD.ZG"), c("gdp"))
head(dtgdp)
tail(dtgdp)
```

```{r GDP_Growth_and_Renewable_Energy_Consumption}
DT8 <- merge(dfrenew,dtgdp , by = "year", all.x = TRUE, all.y = T)
head(DT8)
clabs8 <-   labs(title = "GDP Growth and Renewable Energy Consumption (% of TFEC)",
                subtitle = "Turkey",
                x = "Years", y = "(%)",
                caption = "Source: World Bank")



g20 <- ggplot(DT8, mapping = aes(x = year)) + geom_line(size = 1, mapping = aes(y = renew, color = "renewable energy consumption"))+
  geom_line(size = 1, mapping = aes(y = gdp, color = "GDP Growth")) + my_theme() + guides(col = guide_legend(""))+ clabs8


g21 <-ggplot(DT8, mapping = aes(x = year)) + geom_bar(aes(y = renew, color ="Renewable Energy Cons.(%)"), stat ="identity",position =position_dodge())+
  geom_line(size = 1, mapping = aes(y = gdp, color = "GDP Growth(%)")) + my_theme() +
  guides(col = guide_legend(""))+ clabs8

```

```{r, results=T}
ggplotly(g20)
ggplotly(g21)
```

###  D-Elektrik, gaz, buhar ve iklimlendirme üretimi ve dağıtımı-Düzey Sanayi Üretim Endeksi (2015=100)(TÜİK)
```{r Electricity_gas_steam_and_air_conditioning_supply}
endex <- getDataSeries("TP.SANAYREV4.Y41", CBRTKey= "APA7ENywLk",freq =6, startDate = "01-01-2005", endDate = "01-10-2020",na.rm =T)

class(endex)
head(endex)
setnames(endex, c("time", "TP.SANAYREV4.Y41"), c("year", "index"))

clabs <-   labs(title = "Electricity, gas, steam and air conditioning supply",
                subtitle = "Turkey",
                x = "Years", y = "",
                caption = "Source: CBRT")

g22 <- ggplot(endex, aes(year, index)) + geom_line() + my_theme() + clabs
```

```{r, results=T}
ggplotly(g22)
```

###  Fiyat Endeksi-Özel Kapsamlı TÜFE Göstergeleri (2003=100)(TÜİK)
enerji
```{r Consumer_Price_Index}
energy <-  getDataSeries("TP.FE.OKTG07", CBRTKey= "APA7ENywLk",freq =6, startDate = "01-01-2003", endDate = "01-12-2020",na.rm =T)

class(energy)
head(energy)
setnames(energy, c("time", "TP.FE.OKTG07"), c("year", "energy"))

clabse <-   labs(title = "Consumer Price Index-Indicators for the CPI's Having Specified Coverages (2003=100)",
                subtitle = "Turkey",
                x = "Years", y = "Energy",
                caption = "Source: CBRT")

g23 <- ggplot(energy, aes(year, energy)) + geom_line() + my_theme() + clabse

```

```{r, results=T}
ggplotly(g23)
```

###  gdp growth ile eneergy indexi olmadı bir arada aşağıda görülüyor
```{r gdp_growth_energy_index}
GDPT <- getDataSeries("TP.GSYIH26.HY.ZH", CBRTKey= "APA7ENywLk",freq =6, startDate = "01-01-1998", endDate = "01-07-2020",na.rm =T)
class(GDPT)
head(GDPT)
setnames(GDPT, c("time", "TP.GSYIH26.HY.ZH"), c("year", "GDP"))
head(GDPT)
GDPT[, GDP_Growth := 100*(GDP - shift(GDP, n = 4L))/shift(GDP, n = 4L)]
head(GDPT)

DTenergy <- merge(energy, GDPT, by = "year", all.x = TRUE, all.y = T)
head(DTenergy)

clabs <-   labs(title = "GDP Growth and Energy Index",
                subtitle = "Turkey",
                x = "Years", y = "GDP Growth Rate and Energy Index",
                caption = "Source: CBRT")
g24 <- ggplot(DTenergy, mapping = aes(x = year)) + geom_line(size = 1, mapping = aes(y = energy, color = "Energy Index"))+
  geom_line(size = 1, mapping = aes(y = GDP_Growth, color = "GDP Growth")) + my_theme() + guides(col = guide_legend(""))+ 
  clabs 

```

```{r, results=T}
ggplotly(g24)
```

###  enerji ithalat ihracat ,çeyreklik
Enerji (27. Fasıl) İhracat (Milyon ABD doları)
Enerji (27. Fasıl) İhracat (Milyon ABD doları)-Düzey Cari İşlemler Hesabı, Parasal Olmayan Altın ve Enerji Dış Ticareti (Milyon ABD doları)
```{r energy_import_export}
ihracat <-  getDataSeries("TP.HARICCARIACIK.K5", CBRTKey= "APA7ENywLk",freq =6, startDate = "01-01-1996", endDate = "01-12-2020",na.rm =T)
head(ihracat)
setnames(ihracat, c("time", "TP.HARICCARIACIK.K5"), c("year", "export"))

ithalat <- getDataSeries("TP.HARICCARIACIK.K6", CBRTKey= "APA7ENywLk",freq =6, startDate = "01-01-1996", endDate = "01-12-2020",na.rm =T)
head(ithalat)
setnames(ithalat, c("time", "TP.HARICCARIACIK.K6"), c("year", "import"))

DTimex <- merge(ithalat, ihracat, by = "year", all.x = TRUE, all.y = T)
head(DTimex)

clabsimex <-   labs(title = "Energy Import and Export",
                subtitle = "Turkey",
                x = "Years", y = "",
                caption = "Source: CBRT")
g25 <- ggplot(DTimex, mapping = aes(x = year)) + geom_line(size = 1, mapping = aes(y = import, color = "Energy Import"))+
  geom_line(size = 1, mapping = aes(y = export, color = "Energy Export")) + my_theme() + guides(col = guide_legend(""))+ 
  clabsimex + scale_y_log10()


```

```{r, results=T}
ggplotly(g25)
```

###  yearly data ile de yaptım bu daha iyi oldu
```{r energy_import_export_yearly}
ihracat1 <-  getDataSeries("TP.HARICCARIACIK.K5", CBRTKey= "APA7ENywLk",freq =12, startDate = "01-01-1996", endDate = "01-12-2020",na.rm =T)
head(ihracat1)
setnames(ihracat1, c("time", "TP.HARICCARIACIK.K5"), c("year", "export"))

ithalat1 <- getDataSeries("TP.HARICCARIACIK.K6", CBRTKey= "APA7ENywLk",freq =12, startDate = "01-01-1996", endDate = "01-12-2020",na.rm =T)
head(ithalat1)
setnames(ithalat1, c("time", "TP.HARICCARIACIK.K6"), c("year", "import"))

DTimex1 <- merge(ithalat1, ihracat1, by = "year", all.x = TRUE, all.y = T)
head(DTimex1)


clabsimex1 <-   labs(title = "Energy Import and Export",
                    subtitle = "Turkey",
                    x = "Years", y = "Import",
                    caption = "Source: CBRT")

g26 <- ggplot(DTimex1, mapping = aes(x = year)) + geom_bar(aes(y = import, color = " Import"), stat ="identity",position =position_dodge())+
  geom_line(size = 1, mapping = aes(y = export, color = "Export")) +my_theme() +clabsimex1

```

```{r, results=T}
ggplotly(g26)
```

###  graphs
```{r final_graphs}
aa <-WDIsearch("current account")
bb <-WDIsearch("manufacturing")
cc <-WDIsearch("energy")

cadTR <- WDI(country = "TR", indicator = "BN.CAB.XOKA.CD", start = 1996, end = 2020 )
#Current account balance (BoP)

manufTR <-WDI(country = "TR", indicator = "NV.IND.MANF.CN", start = 1996, end = 2020 )
#Manufacturing, value added (current LCU)

energconsTR <- WDI(country = "TR", indicator = "1.1_TOTAL.FINAL.ENERGY.CONSUM", start = 1996, end = 2020 )
#Total final energy consumption (TFEC)

old1 <- c("iso2c", "country", "BN.CAB.XOKA.CD", "year")
old2 <- c("iso2c", "country", "NV.IND.MANF.CN", "year")
old3 <- c("iso2c", "country", "1.1_TOTAL.FINAL.ENERGY.CONSUM", "year")

setnames(cadTR,old1, c("isocode", "country", "cab", "year"))
setnames(manufTR, old2, c("isocode", "country", "manuf", "year"))
setnames(energconsTR, old3, c("isocode", "country", "energy", "year"))

energconsTR$energy <- round(energconsTR$energy, digits = 3)
cadTR$cab <- cadTR$cab/100000

dt1 <- merge(cadTR, energconsTR, by.x = "year", by.y = "year")

dt2 <- merge(manufTR, energconsTR, by.x = "year", by.y = "year")

#Visualization--------- 

mytheme <- function()  { theme_bw()   + 
      theme(panel.border = element_rect(color = "gray70", size = 1.5)) +
      theme(legend.position = "none") +
      theme(plot.caption = element_text(color = "blue4", size = rel(1))) +
      theme(panel.background = element_rect(fill = "mintcream", colour = "gray70"), panel.grid = element_line(colour ="gray70"), 
            panel.grid.minor = element_line(size = rel(0.5)), 
            strip.background = element_rect(fill = "mintcream", 
                                            colour = "gray70"), complete = TRUE) +theme(axis.title.x= element_text(color = "blue4", size = rel(1.1)),
                                                                                        axis.title.y = element_text(color = "blue4", size = rel(1.1))) + 
      theme(plot.title = element_text(color = "blue4", size = rel(1.5))) + 
      theme(axis.text.x= element_text(color = "blue4",size = rel(0.9)),
            axis.text.y = element_text(color = "blue4",size = rel(0.9))) +
      theme(legend.text = element_text(size = rel(0.8), color = "blue4"))}

labss1 <- labs(title = "Relationship Between Current Account Balance and Energy Consumption",
              x = "Energy Consumption", y = "Current Account Balance (in million)",
              caption = "Source: WorldBank")

labss2 <- labs(title = "Relationship Between Manufacturing and Energy Consumption",
               x = "Energy Consumption", y = "Manufacturing",
               caption = "Source: WorldBank")

bardiag <- ggplot(data = dt1, aes(x=energy, y = cab , color= year )) +geom_point()  +labss1 +mytheme() +
  geom_smooth(method = "lm", se=F, color="purple") +scale_y_continuous(labels = comma)

bardiagg <- ggplot(data = dt2, aes(x=energy, y = manuf , color= year )) +geom_point()  +labss2 +mytheme() +
  geom_smooth(method = "lm", se=F, color="purple") +scale_y_continuous(labels = comma)
```

```{r results=TRUE}
ggplotly(bardiag)
ggplotly(bardiagg)
```

